Reproducible Builds for\\ Computational Research Papers

### 2020-09-16 ms.tex
{make test}

We propose a template for writing computational research papers with \LaTeX\ and verifying the its reproducibility using the concept of \textit{reproducible builds} (RBs) in a local or remote computational environment.
We provide an open source implementation of the template\footnote{\url{https://github.com/pbizopoulos/cookiecutter-reproducible-builds-for-computational-research-papers}} and use it to write this paper\footnote{\url{https://arxiv.org/abs/2005.12660}} and verify its reproducibility\footnote{\url{https://github.com/pbizopoulos/reproducible-builds-for-computational-research-papers}}.
We use Python~\cite{van2007python} for producing the results, however other scientific-oriented programming languages can be used.

For the rest of the paper we will refer to:
\begin{itemize}
	\item \textit{paper} as a computational research paper,
	\item \textit{results} as the figures, tables and variables that are shown in the \textit{paper},
	\item \textit{\LaTeX\ code} as the files containing the main text (\textit{*.tex}) and bibliography (\textit{*.bib}) of the \textit{paper},
	\item \textit{results code} as the files containing the scientific-oriented programming language code (\textit{*.py}) that produces the \textit{results},
	\item \textit{code} as both the \textit{\LaTeX\ code} and \textit{results code},
	\item \textit{author} as the author(s) of the \textit{paper} and
	\item \textit{reviewer} as the reviewer(s) of the \textit{paper} or anyone in interest of verifying the \textit{paper} (author(s), journal editors, other researchers).
\end{itemize}

@@

The proposed template can be applied on \textit{papers} written in \LaTeX~\cite{lamport1994latex} and the \textit{results code} are executed within a Docker container.
\LaTeX\ is a typesetting system for publishing high quality research papers, and Docker~\cite{merkel2014docker} is a container system that allows portable execution of a codebase between different operating systems/environments and has shown great promise in reproducible research~\cite{boettiger2015introduction}.
The corresponding Dockerfile can be provided by the \textit{author} or generated by tools such as repo2docker~\cite{forde2018reproducible}.

A specific file hierarchy is shown in Fig.~\ref{fig:filehierarchyworkflow} however this is optional, as the template only needs to know the directories of the main files of the \LaTeX\ and \textit{results code} during its creation.
The template fits in the research writing and reproducibility verification in the following way:
\begin{enumerate}
	\item the \textit{author} edits the \textit{code} in the local repository,
	\item the \textit{author} may execute \textit{make} during development and experimentation,
	\item the \textit{author} may verify the reproducibility of the \textit{paper} locally by executing \textit{make test} (or remotely by commiting with a message which contains \textit{[make test]}) which triggers the following:
		\begin{enumerate}
			\item the \textit{results code} is executed,
			\item the \textit{results} along with the \textit{\LaTeX\ code} are used to compile a pdf file,
			\item the previous is executed twice,
			\item the \textit{paper} is verifed as reproducible if and only if the pdf files are identical.
		\end{enumerate}
	\item the \textit{reviewer} may verify the reproducibility of the paper locally (by cloning the repo) or remotely (by inspecting the result of \textit{[make test]}).
\end{enumerate}

@@

\subsection{Technical implementation}
We provide an open source implementation of the RBs template in which the remote reproducibility verification is done using Github Actions.
The builder VM uses the latest version of Ubuntu and pulls a docker image for building \LaTeX\ from the Github Container Registry.
We demonstrate its use by writing and verifying the reproducibility of this paper.

Executing RBs can be done using the following:
\begin{lstlisting}[language=Bash, style=lststyle, caption={Makefile call syntax from the shell.}, captionpos=b]
SYNTAX
make [OPTION] [ARGS=--full]

@@

@@

Using the random seeds and \textit{datatool} we could have deterministic stochasticity that reproduces figures, tables and variables.
For example Fig.~\ref{fig:image} depicts the train and validation loss:
\begin{figure*}[!t]
	\subfloat{\includegraphics[width=0.25\textwidth]{results/MNIST-loss}}
	\subfloat{\includegraphics[width=0.25\textwidth]{results/FashionMNIST-loss}}
	\\
	\subfloat{\includegraphics[width=0.25\textwidth]{results/KMNIST-loss}}
	\subfloat{\includegraphics[width=0.25\textwidth]{results/QMNIST-loss}}
	\caption{Train validation losses.
	Image example created from \textit{results code}.}
	\label{fig:image}
\end{figure*}

@@

\section{Discussion}
RBs are based on plain text and are independent of any specific git and docker registry providor thus preventing `vendor lock-in' situations.
Previous works in this area such as Hurlin et al.~\cite{hurlin2019reproducibility} propose an `external certification agency' for economic studies.
RBs provide a simple solution for writing and verifying the reproducibility of a `\textit{paper} as a whole', without having to trust a third-party.
Anyone can clone a repository written with RBs and locally verify its reproducibility, in case third party verifications (such as cloud VMs) cannot be trusted.
Similar to RBs is ReDoc~\cite{schwab2000making} which however focuses on separate result file reproducibility and requires the author to write each rule in the Makefile for each result file.
RBs on the other hand provides a Makefile that is automatically generated by the template.

>>>>

We propose a template for writing computational research papers with \LaTeX\ and verifying  its reproducibility using the concept of \textit{reproducible builds} (RBs) in a local or remote computational environment.
We provide an open source implementation of the template\footnote{\url{https://github.com/pbizopoulos/cookiecutter-reproducible-builds-for-computational-research-papers}}, which we use to write this paper\footnote{\url{https://arxiv.org/abs/2005.12660}} and verify its reproducibility\footnote{\url{https://github.com/pbizopoulos/reproducible-builds-for-computational-research-papers}}.
We use Python~\cite{van2007python} for producing the results, however other scientific-oriented programming languages can be used.

For the rest of the paper we will refer to:
\begin{itemize}
	\item \textit{paper} as a computational research paper,
	\item \textit{results} as the figures, tables and variables that are shown in the \textit{paper},
	\item \textit{\LaTeX\ code} as the files containing the main text (\textit{*.tex}) and bibliography (\textit{*.bib}) of the \textit{paper},
	\item \textit{results code} as the files containing the scientific-oriented programming language code (\textit{*.py}) that produces the \textit{results},
	\item \textit{code} as both the \textit{\LaTeX\ code} and \textit{results code},
	\item \textit{author} as the author(s) of the \textit{paper} and
	\item \textit{reviewer} as the reviewer(s) of the \textit{paper} or anyone interested in verifying the \textit{paper} (author(s), journal editors, other researchers).
\end{itemize}

@@

The proposed template can be used for writing \textit{papers} with \LaTeX~\cite{lamport1994latex} and the \textit{results code} that can be executed within a Docker container.
\LaTeX\ is a typesetting system for publishing high quality research papers, and Docker~\cite{merkel2014docker} is a container system that allows portable execution of a codebase between different operating systems/environments and has shown great promise in reproducible research~\cite{boettiger2015introduction}.

A specific file hierarchy is shown in Fig.~\ref{fig:filehierarchyworkflow} however this is optional, as the template only needs to know the directories of the main files of the \LaTeX\ and \textit{results code} during its creation.
The Makefile and Dockerfile are automatically generated by the template after user prompts.
For more complicated setups the Dockerfile can be provided by the \textit{author} or generated by tools such as repo2docker~\cite{forde2018reproducible}.

The template fits in research writing and reproducibility verification in the following way:
\begin{enumerate}
	\item the \textit{author} edits the \textit{code} in the local repository,
	\item the \textit{author} may execute \textit{make} during development and experimentation,
	\item the \textit{author} may verify the reproducibility of the \textit{paper} locally by executing \textit{make test} (or remotely by commiting the code change with a message which contains \textit{\{make test\}}) which triggers the following:
		\begin{enumerate}
			\item the \textit{results code} is executed,
			\item the \textit{results} along with the \textit{\LaTeX\ code} are used to compile a pdf file,
			\item the previous is executed twice,
			\item the \textit{paper} is verifed as reproducible if and only if the pdf files are identical.
		\end{enumerate}
	\item the \textit{reviewer} may verify the reproducibility of the paper locally (by cloning the repo and executing \textit{make test}) or remotely (by inspecting the result of \textit{\{make test\}}).
\end{enumerate}

@@

\subsection{Technical implementation}
Executing RBs can be done using the following:
\begin{lstlisting}[language=Bash, style=lststyle, caption={Makefile call syntax from the shell.}, captionpos=b]
SYNTAX
make [OPTION] [ARGS=--full]

@@

In this implementation of RBs the remote reproducibility verification is done using Github Actions, the builder VM uses the latest version of Ubuntu and pulls a docker image for building \LaTeX\ from the Github Container Registry.

@@

Using the random seeds and \textit{datatool} we could have deterministic stochasticity that reproduces figures, tables and variables.
For example Fig.~\ref{fig:image} depicts the train and validation loss for all datasets.
\begin{figure}[!t]
	\subfloat{\includegraphics[width=0.5\textwidth]{results/validation-loss}}
	\caption{Validation losses.
	Image example created from \textit{results code}.}
	\label{fig:image}
\end{figure}

@@

\section{Discussion}
RBs are based on plain text files and are independent of any specific version control and docker registry providor thus preventing `vendor lock-in' situations.
Previous works in this area such as Hurlin et al.~\cite{hurlin2019reproducibility} propose an `external certification agency' for economic studies.
RBs provide a simple solution for writing and verifying the reproducibility of a `\textit{paper} as a whole', without having to trust a third-party.
Anyone can clone a repository written with RBs and locally verify its reproducibility, in case third party VMs cannot be trusted.
Similar to RBs is ReDoc~\cite{schwab2000making} which however focuses on separate result file reproducibility and requires the author to write each rule in the Makefile for each result file.
RBs on the other hand provides a Makefile that is automatically generated by the template.
